### 2021-04-26 newvariant.tex
fix

There is currently a (minor) discrepancy in the estimates of relative strength of new variants, particularly for B.1.1.7.
Mathematical analyses have typically reported greater than 1.4 fold increase in reproduction number for the B.1.1.7 variant ($\rho > 1.4$) whereas an independent analysis of secondary attack rates from contact tracing data suggests 1.25--1.4 fold increase \citep{ukinvest}.
We can ask whether different estimates can be reconciled by generation intervals alone---for simplicity, we will assume that the variant and the wild type have identical generation-interval distributions as other studies have.

For example, \cite{davies2021estimated} estimated $\rho=1.77$ assuming a fixed generation generation interval at 5.5 days from $\delta=0.104\pday$ in their regression analysis;
similarly, \cite{volz2021transmission} estimated $\rho=1.79$ by directly comparing the ratio between $\Rv$ and $\Rw$ assuming a mean generation interval of 6.5 days and a squared coefficient of variation of 0.38.
In this case, the latter study's longer mean generation interval (which typically increases $\rho$) is counter-balanced by the effects of their wider distribution, explaining the lack of differences in the estimates of $\rho$ despite the differences in the assumed generation-interval distributions.

However, the fixed generation generation interval assumed by \cite{davies2021estimated} is too narrow.
The mean generation interval of 6.5 days assumed by \cite{volz2021transmission} is also likely too long:
first, this estimate actually corresponds to the observed mean serial intervals from China in between January and February, 2020 \citep{bi2020epidemiology}, during which they were more likely to observe longer serial intervals due to initial epidemic growth \citep{park2021forward, ali2020serial}; and second, recently infected individuals may be less likely to transmit after symptom onset due to increased awareness of the pandemic compared to earlier on, leading to shorter generation intervals.
If we replace the assumed generation-interval distributions for \cite{volz2021transmission} with shorter generation interval estimates from Tianjin, China ($\Gw=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.27$, which is consistent with the attack rate analysis \citep{ukinvest}.
We do not claim that these estimates of $\rho$ or $\Gw$ are more accurate;
instead, this back-of-the-envelope calculation highlights the importance of propagating uncertainty in generation-interval distributions in assessing the relative strength and speed of SARS-CoV-2 variants.

>>>>

There is currently a (minor) discrepancy in the estimates of relative strength of new variants, particularly for B.1.1.7.
Mathematical analyses have typically reported greater than 1.4 fold increase in reproduction number for the B.1.1.7 variant ($\rho > 1.4$) whereas an independent analysis of secondary attack rates from contact tracing data suggests 1.25--1.4 fold increase \citep{ukinvest}.
We can ask whether different estimates can be reconciled by generation intervals alone.

\cite{davies2021estimated} estimated $\rho=1.77$ assuming a fixed generation generation interval at 5.5 days from $\delta=0.104\pday$ in their regression analysis.
Similarly, \cite{volz2021transmission} estimated $\rho=1.79$ by directly comparing the ratio between $\Rv$ and $\Rw$ assuming a mean generation interval of 6.5 days and a squared coefficient of variation of 0.38.
In this case, the latter study's longer mean generation interval (which typically increases $\rho$) is counter-balanced by the effects of their wider distribution, explaining the lack of differences in the estimates of $\rho$ despite the differences in the assumed generation-interval distributions.
We note that both studies performed several analyses to estimate $\rho$ and propagated uncertainty in the assumed generation-interval distributions; these two were chosen for illustrative purposes rather than to question the validity of their estimates.

The fixed generation generation interval assumed by \cite{davies2021estimated} is, however, too narrow.
The mean generation interval of 6.5 days assumed by \cite{volz2021transmission} is also likely too long: 
first, this estimate actually corresponds to the observed mean serial intervals from China in between January and February, 2020 \citep{bi2020epidemiology}, during which they were more likely to observe longer serial intervals due to initial epidemic growth \citep{park2021forward, ali2020serial}; and second, recently infected individuals may be less likely to transmit after symptom onset due to increased awareness of the pandemic, leading to shorter generation intervals.
For example, \cite{davies2021estimated} showed that using shorter generation interval estimates from \cite{ganyani2020estimating} ($\Gw=\Gv=3.6\days$ and $\kappa=0\ \textrm{or}\ 0.7$) can lead to considerable reduction in the estimates of $\rho$ (see Supplementary Table S1 in the original text)---in the case of their regression analysis, $\rho$ decreased from 1.77 to 1.46 (assuming $\kappa = 0$).
Similarly, \cite{volz2021transmission} considered the possibility that the variant may have shorter generation intervals ($\Gv=0.75\Gw$), decreasing $\rho$ from 1.79 to 1.60.
If we consider even shorter generation interval estimates from Tianjin, China ($\Gw=\Gv=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.27$ (from $\delta=0.104\pday$ and $\rw = 0\pday$; \cite{davies2021estimated}).
Although these estimates are more consistent with the attack rate analysis \citep{ukinvest},
we do not claim that these estimates of $\rho$ are more accurate.
Instead, this back-of-the-envelope calculation highlights the importance of propagating uncertainty in generation-interval distributions in assessing the relative strength and speed of SARS-CoV-2 variants.
